Universal Anonymous Signatures

### 2022-03-01 prism-backend/research/gsac/model-uas.tex
Made a pass before sharing.

@@

\commentwho{Jesus}{I'm moving to a syntax that applies arbitrary functions over
  the credentials, at different points (issuance, evaluation, and inspection).
  Concretely, I'm moving away from the selective disclosure type of syntax.
  However, it is my intention that selective disclosure becomes an especial case
  of this. That is, the syntax should support both credentials with plain
  selective attribute disclosure, and ``functional'' credentials.}

@@

\paragraph{Helper Function \Identify.} In addition, we make use of a helper
function \Identify for some of our definitions. This function receives
a signature and a user secret key, and determines whether the latter was used
to produce the former. This is in line with previous works on DAA
\cite{bfg+11,cdl16} and group signatures without traditional open functions
\cite{dl21,fgl21,gl19}. More concretely, the \Identify function, defined as $1/0
\gets \Identify(\usk,\sig)$, conveys meaning to our non-frameability definition
conditioned on the fact that, given a signature \sig accepted by \Verify,
\Identify returns $1$ for exactly one \usk.

\paragraph{Global Variables.} %
The environment manages several global variables in the games posed to the
adversary. Users are referred to with user identifiers, \uid; for credentials,
we use \cid; for groups, \gid. For credentials and groups, we use bold font to
denote sets: i.e., \scid and \sgid denot sets of credential and group
identifiers. All tables/sets are initialized as empty tables/sets, denoted
with $\emptyset$.

@@

\paragraph{Discussion on the generality of anonymity in \UAS schemes.} %
This notion of anonymity is more general than that of group signatures in the
sense that calls to the \INSPECT oracle reveal an arbitrary function of the
identity of the user -- which can certainly be the member index itself, as it
is frequent in group signatures, ore any other function computable from the
user public key, its credentials, and produced signature. Moreover, note that
our notion is even stronger than the conventional ``CCA-like'' anonymity notion
that gives the adversary access to the open oracle, only restricting it when
trying to open challenge signature. Precisely the introduction of generic
inspection functions allows us to \underline{\emph{let the adversary open
    challenge signatures $\csig_b$}} as long as the counterpart $\csig_{1-b}$
makes \Inspect produce the same \y value as output. This is certainly not
possible when using conventional opening, since it directly outputs the identity
of the signer -- which cannot be the same for different challenge users.

@@

\paragraph{Discussion on the generality of traceability in \UAS schemes.} %
The notion of traceability we present for \UAS is strictly more general than
the corresponding one for group signatures. This is again the case as a direct
consequence of the fact that \Inspect, which can return an arbitrary function
of the signer's credential (and signed message), is a generalization of the
conventional \Open -- if we make \Inspect return just the identity of the
signer, then we get something similar to the usual notion of traceability.
Although, even in that case, we need to take into account attributes, and the
fact that the same user may obtain multiple credentials (that is why, even when
having \Inspect return the identity of the signer, our notion is not exactly
the same). And, in this sense, the unforgeability notion that \UAS have is
equivalent to that of anonymous credentials. It would seem, though, that we do
not need the traceability part of group signature; after all, it is the
protection against wrong claims on attributes what enables meaningful and
flexible authentication. However, the type of protection against misuses of
the inspection functionality that we can get with an honest issuer (as in
traceability) is much higher than without an honest user (as in
non-frameability). Specifically, with an honest issuer we can ensure that
the adversary cannot even alter the value returned by \Inspect on signatures
by corrupt users. Whereas, with a corrupt issuer, all we can ensure is that
the adversary cannot forge a signature from an honest user for which \Inspect
returns the same value as a signature by that honest user would produce.
Traceability is, therefore, a core property to ensure accountability
\todo{Discuss this in more detail somewhere else.}

@@

The notion of non-frameability in \UAS schemes is unavoidably more subtle than
in group signatures, though. To see this, we note that, by allowing any
arbitrary inspection function \finsp to be used, it can be perfectly valid to
have a signature produced by a corrupted user, output the same \y value upon
inspection than the one that \finsp outputs when inspecting a signature by an
honest user. As a concrete example, imagine an inspection function that returns
the banking account number (or cryptocurrency address) to be used to request
payment of a fine. In this example, a corrupted user and an honest user could
both have delegated this to some agency\footnote{\todo{Maybe, come up with a
    better example. How about spam filters that apply some sort of classification
    engine? \finsp could be this engine, and emails by many different users could
    be classified just as ``Spam'' or ``Not spam''}}, so it makes perfect sense
that the inspection function returns the same value for both users.
%
Thus, in \UAS schemes, we must be more subtle, by necessity. Note that we cannot
just require that no adversary can produce a signature for which inspection
returns the same value as it would for a signature honestly generated by an
uncorrupted user. This is again the case since inspection may be non-injective,
which means that inspection of a signature by some corrupted user may
legitimately produce the same value as that of an honest user. Thus, we need
to further refine the definition by requiring that (i) no adversary can produce
a signature that is recognized (via \Identify) as originating from an honest
user \uid, (ii) who owns a credential that makes \finsp return a valid \y value
(i.e., generated by \Inspect along with a proof, and accepted by \Judge), (iii)
without having queried the \SIGN oracle before. This non-frameability property
for \UAS schemes is formally defined in \figref{fig:exp-uas-frame}, where
$\Oframe \gets \lbrace\HU,\CU\rbrace\GEN,
\lbrace\II,\OO\rbrace\GEN,\lbrace\II,\OO\rbrace\CORR,\OBTAIN,\SIGN$

\begin{figure}[htp!]
  \procedure{$\ExpNonframe(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
     (\gid,\sig,\msg,\y,\iproof) \gets \adv^{\Oframe}(\parm) \\
     \pcreturn 1~\pcif \Verify(\GK[\gid],\sig,\msg) = 1 \land
     \Judge(\GK[\gid],\y,\iproof,\sig,\msg) = 1~\land \\
     \pcind \pccomment{\adv~did not query \SIGN to get \sig, yet \sig originates
       from an honest \uid, owning a credential that makes \finsp
       evaluate to \y} \\
     \pcind \exists \uid \in \HU~\st~\Identify(\PRVUK[\uid],\sig) = 1 \land
     (\cdot,\sig,\msg,\cdot) \notin \SIG[\uid]~\land\\
     \pcind \exists \cred \in \CRED[\uid]~\st~
     \finsp(\PUBUK[\uid],\cred,\msg) = \y \\
     \pcreturn 0
  }
  \caption{Non-frameability experiment for \UAS schemes.}
  \label{fig:exp-uas-frame}
\end{figure}

@@

\paragraph{Discussion on the generality of non-frameability in \UAS schemes.} %
Anonymous credentials do not have non-frameability property and, thus, it is
hard to make a comparison. However, we can draw some connections with AC schemes
that support revocation, as revocation is somehow equivalent to linking, which
is a type of inspection available in group signatures. In this sense, note that
basic revocation (without straight deanonymization) can be trivially achieved
through our generic \Inspect function. For instance, one could set \finsp to
be a pseudorandom number seeded with the user's public key (or credential). In
this sense, \Inspect could be essentially seen as a Verifiable Random Function.
If we compare with group signatures, our notion is again more general than the
conventional one. Indeed, in typical group signatures, \Open returns the
identity of the signer, and thus does non-frameability not allow the same value
to be returned when opening signatures by different users. Otherwise, a
corrupted user controlled by the adversary could be able to create a signature
opening to some honest user (hence, framing him). But this is again possible
with our definition which, in case of making \Inspect equal \Open, becomes
equivalent to conventional non-frameability. However, it also allows more
generic situations in which signatures by different users may return the same
value.

@@

@@

\todo{Can we leverage the \fissue and/or \feval predicates to make our model
  suitable for some sort of Ring Signature scheme?}

>>>>

We now define our model for Universal Anonymous Signatures. At a very high
level, it is a combination of the main models for group signatures and
anonymous credentials -- we try to extract the most useful features from both.
Instead of considering just one group, as in group signatures, we directly and
explicitly include support for multiple issuers as in anonymous credentials.
Moreover, instead of letting each user own a single credential, we allow
multiple credentials per user, even within the same group. This of course
makes us put forward refined notions of anonymity, traceability and
non-frameability. Note that, property-wise, we keep the three main properties
from group signatures, rather than sticking to the two usual properties of
anonymous credentials (anonymity and unforgeability). This is a consequence
of keeping the inspection capability from group signatures, which we consider
crucial in order to ensure some sort of accountability even in very adversarial
contexts where the issuer may be corrupt.
%
But, as stated in the introduction, we go beyond just combining anonymous
credentials and group signatures, and include support for arbitrary policies
during issuance, signing, and inspection. Consequently, our model needs to
capture the flexibility that these policies provide. In short, manipulations of
the issuance policy are only considered forgeries if the adversary manages to
create a valid signature using a credential that was obtained by dishonestly
manipulating the issuance policy (we discuss this further in the traceability
definition). Signing policies are taken into account both in the anonymity
property, where we need to restrict that both credentials used to produce the
challenge signatures always satisfy the policy, and in traceability.
\commentwho{Jesus}{Think a bit more about the following claim, and update if
  needed.}
We cannot give any assurance on the signing policies when the issuers are
corrupt, and thus they are not considered in the non-frameability property.
Inspection policies are considered both in traceability and non-frameability,
where the adversary wins if it manages to somehow alter the correct output of
\Inspect. What is even suprising, is that given the possible non-injectivity of
the inspection function lets us reach an even stronger notion of anonymity.
Namely, \uline{we can allow the adversary to open challenge signatures, as
  long as the inspection function outputs the same value for both}. We give more
details on these aspects in the actual definitions of the security properties.
%
Finally, from a functional perspective, we require that both issuer and
inspector fix the issuance predicate \fiss and inspection function \finsp when
generating their public keys. Indeed, both will be part of the common group key.
This also makes our definitions easier, as supporting multiple issuance
predicate and inspection functions per group would require to consider many
special cases -- and, anyway, that would probably end up being equivalent to
just requiring that a new group needs to be created per each $(\fissue,\$finsp)$
pair. On the other hand, signing policies can be defined at signing time (in
practice, most probably by the verifiers). This is in line with the usual
practice in anonymous credentials, that let users prove arbitrary claims on
their credentials, as long as they are met by the contained attributes.

@@

@@

\paragraph{Helper Function \Identify.} In addition, we make use of a helper
function \Identify for some of our definitions. This function receives
a signature and a user secret key, and determines whether the latter was used
to produce the former. This is in line with previous works on DAA
\cite{bfg+11,cdl16} and group signatures without traditional open functions
\cite{dl21,fgl21,gl19}. More concretely, the \Identify function, defined as $1/0
\gets \Identify(\usk,\sig)$, conveys meaning to our definitions conditioned on
the fact that, given a signature \sig accepted by \Verify, \Identify returns $1$
for exactly one \usk.

\paragraph{Global Variables.} %
The environment manages several global variables in the games posed to the
adversary. Users are referred to with user identifiers, \uid; for credentials,
we use \cid; for groups, \gid. For credentials and groups, we use bold font to
denote sets: i.e., \scid and \sgid denote sets of credential and group
identifiers. All tables/sets are initialized as empty tables/sets, denoted
with $\emptyset$.

@@

\paragraph{Discussion on the generality of anonymity in \UAS schemes.} %
This notion of anonymity is more general than that of group signatures in the
sense that calls to the \INSPECT oracle reveal an arbitrary function of the
identity of the user -- which can certainly be the member index itself, as it
is frequent in group signatures, or any other function computable from the
user public key, its credentials, and produced signature. Moreover, note that
our notion is even stronger than the conventional ``CCA-like'' anonymity notion
that gives the adversary access to the open oracle, only restricting it when
trying to open challenge signature. Precisely the introduction of generic
inspection functions allows us to \uline{\emph{let the adversary open
    challenge signatures $\csig_b$}} as long as the counterpart $\csig_{1-b}$
makes \Inspect produce the same \y value as output. This is certainly not
possible when using conventional opening, since it directly outputs the identity
of the signer -- which cannot be the same for different challenge users.

@@

\paragraph{Discussion on the generality of traceability in \UAS schemes.} %
The notion of traceability we present for \UAS is strictly more general than
the corresponding one for group signatures. This is again a direct
consequence of the fact that \Inspect, which can return an arbitrary function
of the signer's credential (and signed message), is a generalization of the
conventional \Open -- if we make \Inspect return just the identity of the
signer, then we get something similar to the usual notion of traceability.
Although, even in that case, we need to take into account attributes, and the
fact that the same user may obtain multiple credentials (that is why, even when
having \Inspect return the identity of the signer, our notion is not exactly
the same). And, in this sense, the unforgeability notion that \UAS have is
equivalent to that of anonymous credentials. It would seem, though, that we do
not need the traceability part of group signatures; after all, it is the
protection against wrong claims on attributes what enables meaningful and
flexible authentication. However, the type of protection against misuses of
the inspection functionality that we can get with an honest issuer (as in
traceability) is much higher than without an honest user (as in
non-frameability). Specifically, with an honest issuer we can ensure that
the adversary cannot even alter the value returned by \Inspect on signatures
by corrupt users, nor the output of the signing predicate \feval. Whereas, with
a corrupt issuer, all we can ensure is that
the adversary cannot forge a signature from an honest user for which \Inspect
returns the same value as a signature by that honest user would produce; and,
certainly, a corrupt issuer can arbitrarily issue credentials meeting any
desired predicate \feval .Traceability is, therefore, a core property to ensure
accountability.

@@

The notion of non-frameability in \UAS schemes is unavoidably more subtle than
in group signatures, though. To see this, we note that, by allowing any
arbitrary inspection function \finsp to be used, it can be perfectly valid to
have a signature produced by a corrupted user output the same \y value upon
inspection than the one that \finsp outputs when inspecting a signature by an
honest user. As a concrete example, imagine an inspection function that returns
the banking account number (or cryptocurrency address) to be used to request
payment of a fine. In this example, a corrupted user and an honest user could
both have delegated this to some agency\footnote{\todo{Maybe, come up with a
    better example. How about spam filters that apply some sort of classification
    engine? \finsp could be this engine, and emails by many different users could
    be classified just as ``Spam'' or ``Not spam''}}, so it makes perfect sense
that the inspection function returns the same value for both users.
%
Thus, in \UAS schemes, we must be more subtle, by necessity. Note that we cannot
just require that no adversary can produce a signature for which inspection
returns the same value as it would for a signature honestly generated by an
uncorrupted user. This is again the case since inspection may be non-injective,
which means that inspection of a signature by some corrupted user may
legitimately produce the same value as that of an honest user. Thus, we need
to further refine the definition by requiring that (i) no adversary can produce
a signature that is recognized (via \Identify) as originating from an honest
user \uid, (ii) who owns a credential that makes \finsp return a valid \y value
(i.e., generated by \Inspect along with a proof, and accepted by \Judge), (iii)
without having queried the \SIGN oracle before. This non-frameability property
for \UAS schemes is formally defined in \figref{fig:exp-uas-frame}, where
$\Oframe \gets \lbrace\HU,\CU\rbrace\GEN,
\lbrace\II,\OO\rbrace\GEN,\lbrace\II,\OO\rbrace\CORR,\OBTAIN,\SIGN$

\begin{figure}[htp!]
  \procedure{$\ExpNonframe(1^\secpar)$}{%
    \parm \gets \Setup(1^\secpar) \\
     (\gid,\sig,\msg,\y,\iproof) \gets \adv^{\Oframe}(\parm) \\
     \pcreturn 1~\pcif \Verify(\GK[\gid],\sig,\msg) = 1 \land
     \Judge(\GK[\gid],\y,\iproof,\sig,\msg) = 1~\land \\
     \pcind \pccomment{\adv~did not query \SIGN to get \sig, yet \sig originates
       from an honest \uid, owning a credential that makes \finsp
       evaluate to \y} \\
     \pcind \exists \uid \in \HU~\st~\Identify(\PRVUK[\uid],\sig) = 1 \land
     (\cdot,\sig,\msg,\cdot) \notin \SIG[\uid]~\land\\
     \pcind \exists \cred \in \CRED[\uid]~\st~
     \finsp(\PUBUK[\uid],\cred,\msg) = \y \\
     \pccomment{\todo{I don't think we can give any assurance about \feval here,
         given that issuers are corrupt. But still, think about it.}} \\
     \pcreturn 0
  }
  \caption{Non-frameability experiment for \UAS schemes.}
  \label{fig:exp-uas-frame}
\end{figure}

@@

\paragraph{Discussion on the generality of non-frameability in \UAS schemes.} %
Anonymous credentials do not have non-frameability property and, thus, it is
hard to make a comparison. However, we can draw some connections with AC schemes
that support revocation, as revocation is somehow equivalent to linking, which
is a type of inspection available in group signatures. In this sense, note that
basic revocation (without straight deanonymization) can be trivially achieved
through our generic \Inspect function. For instance, one could set \finsp to
be a pseudorandom number seeded with the user's public key (or credential). In
this sense, \Inspect could be essentially seen as a Verifiable Random Function.
If we compare with group signatures, our notion is again more general than the
conventional one. Indeed, in typical group signatures, \Open returns the
identity of the signer, and thus non-frameability does not allow the same value
to be returned when opening signatures by different users. Otherwise, a
corrupted user controlled by the adversary could be able to create a signature
opening to some honest user (hence, framing him). But this is again possible
with our definition which, in case of making \Inspect equal \Open, becomes
equivalent to conventional non-frameability. However, it also allows more
generic situations in which signatures by different users may return the same
value.

@@

\commentwho{Jesus}{This section is aimed at showing the ``universality'' of
  our model. If I'm not mistaken, ideally, it should support the following
  variations (except the transformation to interactive presentations,
  probably).}

@@

\todo{Can we come up with a (\fissue,\feval,\finsp) tuple that allows us to
  somehow mimic ring signatures (i.e., ``a group signature without issuer nor
  opening''? Does the model support this?}
